---
title: "Git_In_R"
author: "David"
date: "2023-10-16"
output: html_document
---

## HOLA! 

So, here we are! We are Monday 23 (i hope) and we are late in the afternoon and tired about the whole traveling... but! Happy and excited to meet and begin this 3 days trip through ecology!!! 

But... who says "nice trip" through ecology might also mean... "numerical ecology" or "using R to later see ecological stuff"... Yes, I agree that it is always nicer to see the wonders that remain hidden in our ecosystems and try to guess how do they function...how this small animal that has milions and milions of years of evolution on his DNA has arrived at that specific site in that specific moment and now is thriving and welcoming you, ready to discover and protect it of the terrible things that us are doing to it, you look at its eyes... it looks you back and... *WAKE UP!!!*

Stop dreaming on bullshit! This is the real world! You are going to put that bug in alcohol to later look at it in the lab and then submerge yourself in R to try to understand WTF is going on!! 

SO! IT IS ALREADY TIME TO ASSUME THAT IT WAS NOT AS *David Attenborough* showed to you!! THIS IS NOT THE BBC! WE ARE IN THE FRONT LINE AND WE WANT TO SAVE THIS SMALL AND DISGUSTING THINGS FROM EXTINCTION!!! SO.... let's use R :O) 

Or at least, let's make our lives much more easier with R because at the end it is not 'that big deal' 

SO, let's begin by the basics of the basics
https://jennybc.github.io/2014-05-12-ubc/ubc-r/session03_git.html


```{r In R we trust }

#### THE PROGRAM

# Basic concepts of the program
# The script 
# Console
# Environemnt

# Files - Plots - Packages - Help 
# Internet - stackoverflow etc etc etc  

# THE PHILOSOPHY!!!!!!!

### The script and how to operate
# Scripts
## How to write stuff 

Hola
"Hola"
#Hola 
()
{}
[]

3
3
3
3

(3+5)/5

# Search  
HOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLA
HOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLA
HOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAADEUHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLAHOLA


# R BASICS - types of objects (most used ones)
### https://rstudio-education.github.io/hopr/r-objects.html 

## Vector
c()

# Numeric 
c(1,2,3,4,5)

# Character
c("Hola","Adeu", "Benvingut")

#Logical 
c(TRUE, FALSE, TRUE)

## Matrix 
matrix()

# --- "FUNCTIONS"!
matrix()

#HELP!
#In the script
?matrix()
# Manually in the help search 
# Using tab when writing the function 

matrix(1,2,3)

matrix(
  c(1,2,3,4,5),
  nrow =2, ncol = 5
  )

matrix(
  c(1,2,3,4,5),
  nrow =5, ncol = 2
)

# Lists 

list(
  c(1,3,4,5,6,6),
  c("Manolo","Paco","Lorenzo"),
  "R és toh guapo",
  c(TRUE, FALSE, TRUE, TRUE)
)

# Data.frame 
# https://rstudio-education.github.io/hopr/r-objects.html#data-frames

data.frame(  c(1,3,4,5,6,6),
             c("Manolo","Paco","Lorenzo"),
             "R és toh guapo",
             c(TRUE, FALSE, TRUE, TRUE)
)
# Same lenghts
data.frame(  c(1,3,4),
             c("Manolo","Paco","Lorenzo"),
             c(TRUE, FALSE, TRUE)
)

# Variables names
data.frame( money= c(1,3,4),
            names=  c("Manolo","Paco","Lorenzo"),
            thief=  c(TRUE, FALSE, TRUE)
)

# Matrices vs DataFrames
# https://www.geeksforgeeks.org/matrix-vs-dataframe-in-r/ Matrix vs DataFrame

### Interact with objects 

c(1,2,3,5)

a <- c(1,2,3,5)
a

pots_posarli_el_nom_el_que_vulguis <- c(1,2,3,5)
pots_posarli_el_nom_el_que_vulguis

length(pots_posarli_el_nom_el_que_vulguis)

pots_posarli_el_nom_el_que_vulguis[5]

pots_posarli_el_nom_el_que_vulguis[5] <- 3



el_data_frame_que_teniem_abans <- data.frame( money= c(1,3,4,6),
                                              names=  c("Manolo","Paco","Lorenzo","Joan Antoni"),
                                              thief=  c(TRUE, FALSE, TRUE,TRUE)
)
el_data_frame_que_teniem_abans

el_data_frame_que_teniem_abans$money
el_data_frame_que_teniem_abans$names

length(el_data_frame_que_teniem_abans)

colnames(el_data_frame_que_teniem_abans)

nrow(el_data_frame_que_teniem_abans)
ncol(el_data_frame_que_teniem_abans)



my_super_cool_field_dataset <- data.frame(
                                          face = c("king", "queen", "jack", "ten", "nine", "eight", "seven", "six",
                                                   "five", "four", "three", "two", "ace", "king", "queen", "jack", "ten", 
                                                   "nine", "eight", "seven", "six", "five", "four", "three", "two", "ace", 
                                                   "king", "queen", "jack", "ten", "nine", "eight", "seven", "six", "five", 
                                                   "four", "three", "two", "ace", "king", "queen", "jack", "ten", "nine", 
                                                   "eight", "seven", "six", "five", "four", "three", "two", "ace"),  
                                          suit = c("spades", "spades", "spades", "spades", "spades", "spades", 
                                                   "spades", "spades", "spades", "spades", "spades", "spades", "spades", 
                                                   "clubs", "clubs", "clubs", "clubs", "clubs", "clubs", "clubs", "clubs", 
                                                   "clubs", "clubs", "clubs", "clubs", "clubs", "diamonds", "diamonds", 
                                                   "diamonds", "diamonds", "diamonds", "diamonds", "diamonds", "diamonds", 
                                                   "diamonds", "diamonds", "diamonds", "diamonds", "diamonds", "hearts", 
                                                   "hearts", "hearts", "hearts", "hearts", "hearts", "hearts", "hearts", 
                                                   "hearts", "hearts", "hearts", "hearts", "hearts"), 
                                          value = c(13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3, 2, 1, 13, 12, 11, 10, 9, 8, 
                                                    7, 6, 5, 4, 3, 2, 1, 13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3, 2, 1, 13, 12, 11, 
                                                    10, 9, 8, 7, 6, 5, 4, 3, 2, 1)
)

my_super_cool_field_dataset

# How many samples (rows) does it have?

### LOAD DATA

# .RData files 

# LOOK AT THE CONSOLE!
load("C:/Users/David CM/Dropbox/DAVID DOC/LLAM al DIA/1. FEHM coses al DIA/6. Infinite R/1.load_data.RData")

# Many ways to load data! 

# Direct link 
load("bla/bla/bla/bla/bla/bla/bla/bla/bla/THE_NAME_OF_THE_THING.RData")

# Set working directory
# - Manually 
# - Script

getwd()
setwd("C:/Users/bla/bla/bla/bla/bla/bla/bla")

setwd("C:/Users/David CM/Dropbox/DAVID DOC/LLAM al DIA/1. FEHM coses al DIA/6. Infinite R")

load("1.load_data.RData")

# Managing datasets from excel to R

# First - Folder with the data
# PROJECTS

getwd()

# Manually 
# Inside the script
library(readxl)
MyData <- read_excel("Folder with the data/MyData.xlsx",
                     col_types = c("text", "text", "text","numeric","numeric"))
View(MyData)
MyData

# .csv (more stable)
MyData <- read.csv2("Folder with the data/MyData.csv")
str(MyData)

MyData <- read.csv2("Folder with the data/MyData.csv",sep = ",",dec = ".")
str(MyData)

MyData <- read.csv2("Folder with the data/MyData.csv", sep = ";",dec = ".",
                    col.names = c("Site","Wild","Taxa","Abun_L"))
str(MyData)

MyData <- read.csv2("Folder with the data/MyData.csv", sep = ";",dec = ".")
str(MyData)

colnames(MyData)
colnames(MyData) <- c("Site","Wild","Taxa","Abun_L")
str(MyData)

MyData

is.matrix(MyData)
is.data.frame(MyData)

# Last things before starting to "dig into it"

### How to order stuff in R

# Use headers! 
# This is the title of what I will do below ####

# Another title of another analysis ####

# 1. Another title of another analysis ####

#____________________________________________________________________________ lines 

#____________________________________FINAL__________________________________####
```

But R goes beyond all this and there are packages that make our lives much more easier

```{r Weolcom to DPLYR!}
# Write next to the functions that you use, things about how you do it, why, whatever
MyData <- read.csv2("Session_0_RRRedondeaRRR/Folder with the data/MyData.csv", sep = ";",dec = ".",header = F) # Here I load the data 
colnames(MyData) <- MyData[1,]
MyData <- MyData[-1,]
# IMPORTANT: decimal is "."! 
# Write down everything!!!!! Key for understanding what you did after 2 days have passed ;)  

### The functions or operations
## Load packages

# Manually
# Inside the script
install.packages("babynames")
library(babynames)
# CRAN or websites for more info or even retrieve directly from GITHUB or similars


### Data exploration
extinction <- read.csv2("Session_0_RRRedondeaRRR/Data explor/extinction.csv", sep = ",",dec = ".")

summary(extinction)

plot(extinction)
plot(extinction, col="red")
str(extinction)

head(extinction)

hist(extinction$body_size)

boxplot(extinction$extinction_risck, col="red")

pairs(extinction)

cor(extinction,method = "spearman")

scatter.smooth(extinction, col="blue")

pairs(extinction ,lower.panel = panel.smooth)
corrmorant::corrmorant(extinction)

```


```{r Still welcome to DPLYR}

# One other way
install.packages("readr")
library(readr)

extinction <- read_csv("Data explor/extinction.csv")
?readr

extinction

# Deactivate the package 
# Manually
#In the script
detach("package:readr", unload = TRUE)

### Data management (one of the ways to do it)

install.packages("tidyvierse")

library(tidyverse)
library(dplyr)

# Working with dplyr package allows to work with data.frames and modify/change datasets inside R 

# DATA MUST BE IN A TIDY FORMAT !!!! (rows = samples) (colums = variables/others)

# filter
# select
# arrange
# mutate
# summarise

# All these comands are actions that can be performed in dplyr
# First le's explore the data  
library(babynames)

data(babynames)
babynames
?babynames

summary(babynames)

plot(babynames$year, babynames$n)

boxplot(babynames$n~babynames$sex)

boxplot(log(babynames$n)~babynames$sex, col=c("red", "blue"))

# Locating some values?
max(babynames$n)
min(babynames$n)

which(babynames$n==max(babynames$n))

babynames[which(babynames$n==max(babynames$n)),]

# DPLYR functions are an easier setting to manage these data
# Step by step, function by function of dplyr

# FILTER
#old way to use the same filter
babynames[babynames$name=="Alice",]

# with tidyverse
filter(babynames, name== "Alice", sex=="F")

filter(babynames, year>2016)

filter(babynames, name %in% c("Ana", "Alice"))

filter(babynames, sex=="F", prop>0.07)

# SELECT

select(babynames, year)

select(babynames, -year)

select(babynames, name, sex)

select(babynames, 1:3)

select(babynames, starts_with('n'))

# ARRANGE

arrange(babynames, prop)

arrange(babynames, desc(prop)) # asc() is the default

arrange(babynames, desc(name), desc(n))

# MUTATE

mutate(babynames, total=n/prop)

mutate(babynames, prop = prop*100)

mutate(babynames, year_diff=2018-year, months_diff=year_diff-12)

mutate(
  babynames,
  dummy=if_else(sex=="F", "Femenino", "Masculino")# if_else IS BINNARY
)

mutate(
  babynames,
  dummy = case_when(
    name == "Alice" ~ "Alicia",
    name == "Anna" ~ "Anne",
    name == "Emma" ~ "Emmmmah",
    TRUE ~ "pues otro nombre"
  )
)

# SUMMARISE 

summarise(
  babynames, 
  mean= mean(year),
  sd= sd(year)
)

# GROUP_BY

by_year <- group_by(babynames, year)

summarise(by_year, mean= mean(n))

plot(summarise(by_year, mean= mean(n)))


# Piping!


babynames%>%
        group_by(year)%>%
        summarise(mean= mean(n))

# Pipes connect each call that you do in "dplyr" is like saying 
#     And then --- 
#     And then --- 

# Not using pipes
summarise(
           group_by(
           filter(babynames, year == 1880), sex
                   ),
            max = max(n),prop = max(prop))

# Using them 

year_1880 <- babynames %>%
                          filter(year == 1880) %>%
                          group_by(sex) %>%
                          summarise(max = max(n),prop = max(prop))


babynames%>%
  group_by(year, sex)%>%
  arrange(desc(prop))%>%
  summarise(max= max(prop),
            name=first(name)) %>%
  select(name, sex, year, max)


### PLOT IN THE SAME ENVIRONMENT! 

install.packages("ggplot2")
library(ggplot2)

years <- babynames %>%
  filter(year == c(1880,1910,1950,1960)) %>%
           group_by(year, sex) %>%
           summarise(mean = mean(n))

seq(1880, 2017, 10)

years <- babynames %>%
  filter(year == seq(1880, 2017, 10))%>%
  group_by(year, sex) %>%
  summarise(mean = mean(n))

ggplot(data=years,
       aes(x = year, y = mean))+
  geom_point()
  

# The final value!!!!
  
babynames%>%
            filter(year==seq(1880, 2017, 10))%>%
            group_by(year, sex)%>%
            summarise(mean = mean(n))%>%
            ggplot(
                   aes(x = year, y = mean, color=sex))+
                   geom_point(size=5)+
                   geom_line()+
                   scale_color_manual(values=c("pink","blue"))+
                   theme_classic()

```

```{r Going WILD}
#####__________________________________________________________________________________
## TIME TO GO WILD! 

MyData <- read.csv2("Session_0_RRRedondeaRRR/Folder with the data/MyData.csv",dec = ".")
colnames(MyData) <- c("ID", "Wildfire","Taxa","AbundL","Time")
str(MyData)
head(MyData)

Matriu_Ab_Spp <- MyData%>%
                    group_by(Wildfire,Time,Taxa)%>%
                    summarise(TotAbb=sum(AbundL))

MyData%>%
  group_by(Time, Wildfire,Taxa)%>%
  summarise(TotAbb=sum(AbundL))%>%
  mutate(PA=1)%>%
  group_by(Time,Wildfire)%>%
  summarise(Rich=sum(PA))%>%
  ggplot(aes(x=Time, y=Rich, fill=Wildfire))+
        geom_line(aes(col=Wildfire),size=2)+
        geom_point(size=5, shape=21)+
  scale_x_discrete(breaks=c("Spring","Summer","Autum","Winter"),
                    limits=c("Spring","Summer","Autum","Winter"))+
  theme_classic()

install.packages("vegan")
library(vegan)
install.packages("adespatial")
library(adespatial)

?specnumber

Matriu_SitxSpp <- Matriu_Ab_Spp%>%
                  spread(key=Taxa, value = TotAbb, fill = 0)

#Richness
richne <-specnumber(Matriu_SitxSpp,MARGIN = 1)

#Beta-diversity partitions
taxa.q_Jaccard <- beta.div.comp(Matriu_SitxSpp[3:ncol(Matriu_SitxSpp)],
                                coef = "J", quant = TRUE)
taxa.q_Jaccard$D
taxa.q_Jaccard$repl
taxa.q_Jaccard$rich

#Beta-diversity - Uniqueness
LCBD_values <- beta.div(Matriu_SitxSpp[3:ncol(Matriu_SitxSpp)], method = "hellinger", nperm = 9999)
LCBD_values$LCBD

# Multivariance NMDS

NMDS_values <- metaMDS(Matriu_SitxSpp[3:ncol(Matriu_SitxSpp)], distance = "bray")

plot(NMDS_values,display = "sites")

x<- NMDS_values$points[,1]
y<- NMDS_values$points[,2]

dataset <- data.frame(x,y,
                      Wildfire=Matriu_SitxSpp$Wildfire)
dataset%>%
ggplot(aes(x=x, y=y, fill=Wildfire))+
           geom_point(shape=21, size=6)+
  theme_classic()

```






